\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-06 pldi/draft.tex
Checkpoint

%% Author with single affiliation.
\author{
  Anonymous Author(s)
  \normalsize{Additional Anonymized Material: \href{https://github.com/u-3/dish}{https://github.com/u-3/dish}.}
}
%% \authornote{with author1 note}          %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

%% \kk{Decide if we will mainly focus on scripts or pipelines and
%%   choose wordings appropriately.}
%
Consider the following data processing Shell script.
%% \kk{We might want to have a more complex script here.}

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
cat * | tr -cs A-Za-z\n | tr A-Z a-z |    ($$(p_1)$$)
  sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}

\noindent
Scripts like the one above are commonly used for orchestrating
data-processing workflows; connecting system commands with custom
tools to perform complex tasks.
%% \kk{Do we need to mention domains where that happens?}
Even though shell pipelines are designed to exploit task parallelism,
they completely ignore available data-parallelism. For small input
sizes this isn't a problem since scripts take less than a second to
execute, but as data sizes increase, failing to utilize the underlying
parallel computational resources is a significant bottleneck.

To address this concern, a Shell \emph{user} could resort to several
different solutions (e.g. gnuparallel, tsc, flags in individual
commands etc...) that offer some form of parallelism. However these
solutions are too ad-hoc (each having a disparate parallelism model),
too limited (they only apply to specific use-cases), and too
coarse-grained. \kk{Tone the previous claims down and back them up.}
At the same time, a \emph{developer} of Shell commands could rewrite
her commands using low level tools (e.g. Raftlib, openMP, etc) to
improve utilization and performance. Manual rewriting is expensive and
can introduce new bugs, cascading changes, and divergence from legacy
functionality.

To address this issue for both \emph{users} and \emph{developers}, we
propose \sys, a tool that automatically exploits available
data-parallelism in a given shell script. The key insight behind \sys
is that a large subset of the Shell language can be encoded as a
dataflow graph, a flexible representation that allows for
parallelization optimizations. \sys builds on this insight with a
careful study of the parallelizability properties of shell primitives
and commands, a definition of a dataflow graph model that is suitable
for the shell, a set of parallelization optimizations that preserve
the behaviour of the graph, and a late-bound command-prefixing scheme
for deferring optimizations as late-as-possible (\ie when runtime
information is available). \kk{Maybe elaborate some more on the
  contributions in this paragraph.}  A \sys-enabled $p_1$ will run
\ttt{tr}s on parallel streams, use a mostly-parallel tree of
\ttt{sort}s, run \ttt{uniq} mostly in parallel, execute \ttt{head} in
one of the streams, and (trivially) merge streams in \ttt{out}. The
only requirement to support a new command is that its developer adds a
lightweight annotation to describe its parallelizability properties.

The paper is structured as follows. It starts by introducing the
necessary background on shell scripting and overviewing \sys~\sx{bg}.
Sections \ref{distributability}--\ref{impl} highlight key
contributions:
\begin{itemize}
  \item
  \S\ref{distributability} studies the parallelizability properties of
  a large class of Shell commands (coreutils) and introduces an
  lightweight annotation language for commands that could be executed
  in a data-parallel manner. \kk{Mention that we did this for all
    coreutils.}

  \item
  \S\ref{ir} presents a dataflow graph model for encoding shell
  scripts, and a set of safe parallelization transformations that
  preserve the semantics of the sequential program.

  \item
  \S\ref{impl} describes the \sys implementation and addresses several
  challenges related to the translation and optimization of shell
  programs. \kk{Rewrite that when we know for sure what the section
    contains.}

@@

\noindent
\sys's evaluation~\sx{eval} shows significant speedups (4--10$\times$)
for parallel execution using a combination of real pipelines.
%micro-bench\-marks %---popular shell one-liners that highlight
%certain features---and a multi-line macro-benchmark---a popular
%weather analysis script.
After a comparison with related prior work~\sx{related}, the paper
closes with a discussion of limitations and possible future
directions~\sx{discussion}.
%% Additional material, in anonymized form, is available at
%% \href{https://git.io/Je6Nk}{https://git.io/Je6Nk}.

% TODO: Important to have a small script that highlights
\nv{
(i) different classes of commands
(ii) challenge of shell expansion order
(iii) segments that are parallelized by \sys (with provable correctness)
(iv) segments that are \emph{not} parallelized by \sys (otherwise they can lead to
incorrectness)
}

@@

As this is benchmark is taken directly from the Hadoop book, it is worth reporting on its execution time.
The comparison is not unreasonable as both systems
  (i) favor message-passing rather than shared memory,
	(ii) Hadoop has no fault-tolerance for master and worker nodes (\ie both systems do not replicate data),
	(iii) the task at hand was presumably chosen by the book author as one that matches Hadoop's capabilities, and 
	(iv) the Hadoop program is written as a single program tailored for this task---with ample opportunity for optimization from the compiler and runtime system---rather than a series of loosely coupled programs that each written to handle many other cases.
\todo{The original pipeline executes in XXs}.
With two Hadoop 3.2.1 (Openjdk 11.0.5) data-nodes, the Hadoop program takes 6m22s for data processing and 1m09s for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---this includes downloading and uncompressing the data files which, as described earlier, is not expressible in the Hadoop equivalent.

>>>>

%% Author with single affiliation.
\author{
  Anonymous Author(s)\\
  \normalsize{Additional Anonymized Material: \href{https://git.io/Je6Nk}{https://git.io/Je6Nk}.}
}
%% \authornote{with author1 note}          %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

%% \kk{Decide if we will mainly focus on scripts or pipelines and
%%   choose wordings appropriately.}
%
% Consider the following data processing Shell script.
% %% \kk{We might want to have a more complex script here.}
% 
% \begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
% cat * | tr -cs A-Za-z\n | tr A-Z a-z |    ($$(p_1)$$)
%   sort | uniq -c | sort -rn | head 5 > out
% \end{lstlisting}
% 
% \noindent
% great sentence -- keep
% Scripts like the one above are commonly used for orchestrating data-processing workflows; connecting system commands with custom tools to perform complex tasks.
% %% \kk{Do we need to mention domains where that happens?}
% Even though shell pipelines are designed to exploit task parallelism, they completely ignore available data-parallelism.
% % Too low-level descriptions of the problem, maybe keep it high level?
% For small input sizes this isn't a problem since scripts take less than a second to execute, but as data sizes increase, failing to utilize the underlying parallel computational resources is a significant bottleneck.

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language agnosticism, coupled with \unix's toolbox philosophy~\cite{mcIlroymemo}, makes the shell the primary choice for specifying succinct and simple pipelines that target data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort and is shared unevenly between two different groups (or roles) of a programmer. % it's more like personas or "wearing hats"

The first group is \emph{command developers} responsible for developing individual commands such as \ttt{sort}, \ttt{uniq}, and \ttt{jq}.
These developers usually work in a single---often low-level---programming language and manually incorporate parallelism through means available in their chosen language, whenever this is possible. %  by the problem and language.
% the language and abstractions they are using focused at a particular task 
% also: target a small problem but aim at high reusability 
As they have no visibility into the command's uses, they expose a plethora of ad-hoc command-specific flags such as \ttt{-t}~\cite{}, \ttt{--parallel}~\cite{}, \ttt{-p}, \ttt{CPUs}, and \ttt{NUM\_THREADS}.

The second group is \emph{command users}, composing scripts out of individual commands.
% Unfortunately, users do not operate in a single language
As their programs combine a variety of commands written in many languages, they are left with only a few manual options for achieving parallelism:
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
either coarse-grained tools such as \eg GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, or manual parallelism POSIX shell primitives and 
% (ii) ad-hoc flags provided by command developers and mentioned above.

Their last resort is 
and writing the program from scratch
Manual rewriting is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
Often, these users are domain experts lacking a deep understanding of parallel systems---in desperate need of automation when it comes to parallelization.

This paper presents a solution targeting the second class of programmers, command users, 
\sys can automate

A system that is aware of the \unix philosophy, its key abstractions, and offers automated POSIX-aware parallelization as long as 
These lightweight annotations can be expressed by developers 
can be expressed 
This is paired with parallelizability study that informed the design of the system and its annotation interface, and led to the parallelizability semantics for POSIX and GNU coreutils---a large class of commands that serves as the standard library.

%% First, shell users composing pipelines (or simply running legacy pipelines on massive datasets) can see scalability benefits without any manual effort---no need for \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}, or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.

To address this challenge, we develop \sys, a system for transforming POSIX shell scripts to their parallel equivalents.
\sys packs several ideas around the insight that a large subset of the Shell language can be represented as a dataflow graph, enabling for parallelization optimizations.
First, a careful study of the parallelizability properties of shell primitives and commands.
\sys addresses the challenge by offering command developers a unified set of abstractions for describing the parallelizability characteristics of their commands.
Using lightweight annotations akin to type signatures, \sys can automate the parallelization of entire POSIX scripts without any input from their user.
hand-in-hand with the parallelizability study:
  the study informed the annotation language and was used to annotate the entire
  for all 
---a set of annotations that \sys already offers for 
for the standard library
  akin to the standard library
  the annotation language is informed by the study, but also the study is 
XX to a type signature, these annotations describe

Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
(one sentence for the model)
(one sentence for the optimizations)
The result of the compilation is still a POSIX-compliant script that manages parallelism explicitly;
  POSIX compliance means 
it does not depend on a managed runtime

Third, \sys packs a runtime component 
A small library of highly-optimized data combiners allows 
This is paired with a
it comes with a few runtime primitives for eager data splitting and merging that address several practical issues ranging from deadlocks, performance bottlenecks;
  these primitives were developed from uncovering several pathological situations, on many of which we report.

We perform an extensive evaluation of \sys using 
  (i) a series of benchmarks, ranging from classic scripts found in the \unix literature to modern data-processing pipelines;
  (ii) three large and complex use cases, ranging from temperature analysis to web crawling and indexing; and
  (iii) a series of micro-benchmarks inspired by (and comparing with) possible alternatives.
The results demonstrate significant advantages in correctness, manual effort, and performance---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Improvements on the performance of individual commands depend on several factors such as the script's execution time, the pipeline depth and the parallelizability class under which commands fall.

%                                           a tool that automatically exploits available data-parallelism in a given shell script.

The paper is structured as follows.
It starts by introducing the necessary background on shell scripting and overviewing \sys~\sx{bg}.
Sections \ref{distributability}--\ref{impl} highlight key contributions:
\begin{itemize}
  \item
  \S\ref{distributability}
    studies the parallelizability properties of a large class of Shell commands and introduces a lightweight annotation language for commands that can be executed in a data-parallel manner.

  \item
  \S\ref{ir} presents a dataflow graph model for encoding shell scripts, and a set of parallelization transformations that focus on preserving the semantics of the sequential program.

  \item
  \S\ref{impl} describes the \sys implementation and addresses several challenges related to the translation and optimization of shell programs.
% \kk{Rewrite that when we know for sure what the section contains.}

@@

\noindent
After \sys's evaluation~\sx{eval} and  comparison with related work~\sx{related}, the paper closes with a discussion of limitations and possible future directions~\sx{discussion}.
%micro-bench\-marks %---popular shell one-liners that highlight
%certain features---and a multi-line macro-benchmark---a popular
%weather analysis script.
%% Additional material, in anonymized form, is available at
%% \href{https://git.io/Je6Nk}{https://git.io/Je6Nk}.

% % TODO: Important to have a small script that highlights
% \nv{
% (i) different classes of commands
% (ii) challenge of shell expansion order
% (iii) segments that are parallelized by \sys (with provable correctness)
% (iv) segments that are \emph{not} parallelized by \sys (otherwise they can lead to
% incorrectness)
% }

@@

As this is benchmark is taken directly from the Hadoop book, it is worth reporting on its execution time.
While the setup is not typical of Hadoop, the comparison between the two systems is not unreasonable:  % we took several steps to make the comparison fair:
  (i) both Hadoop and Spark favor message-passing rather than shared memory,
	(ii) Hadoop has no fault-tolerance for master and worker nodes (\ie both systems do not replicate data),
	(iii) the task at hand is a good fit for  Hadoop,
	(iv) the Hadoop program is written as a single program tailored for this task---with ample opportunity for optimization from the compiler and runtime system---rather than a series of loosely coupled programs each written to handle many other cases.
\todo{The original pipeline executes in XXs---long enough to amortize Hadoop's 0.5--0.9s startup costs.}
With two Hadoop 3.2.1 (Openjdk 11.0.5) data-nodes, the Hadoop program takes 6m22s for data processing and 1m09s for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---this includes downloading and uncompressing the data files which, as described earlier, is not expressible in the Hadoop equivalent.
